---
title: "Table One for COG merged MLL/NSD1 cohort slices"
author: "Timothy J. Triche, Jr. (on behalf of all authors)"
date: "2023-2-1"
output:
  html_document:
    fig_width: 8
    code_folding: hide
    highlight: tango
    toc: yes
    keep_md: true
    toc_float:
      collapsed: yes
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=4,
                      echo=FALSE, warning=FALSE, message=FALSE)
# rmarkdown::render("tableOne.Rmd", "pdf_document") # to render as PDF
```

# Covariates and outcomes for MLL and NSD1 fusion patients

Patient number 368 does not have an annotated sex or age in the COG CDEs. 

```{r}
require(bifurcatoR)
data(dxsmall, package="bifurcatoR")
covs <- as.data.frame(colData(dxsmall))
```

## Aggregate 

Tabulate age group, sex, protocol, and outcome overall.
NA will be tabulated whenever present. 

```{r}
if (requireNamespace("tableone")) {
  CreateTableOne(data=covs, includeNA=TRUE)
}
```

## Per trial

Tabulate age group, sex, and outcome per trial  

```{r}
if (requireNamespace("tableone")) {
  CreateTableOne(data=subset(covs), strata="Protocol", includeNA=TRUE)
}
```


## By fusion

Tabulate age group, sex, protocol, and outcome by fusion group

```{r}
if (requireNamespace("tableone")) {
  CreateTableOne(data=subset(covs), strata="FusionGroup", includeNA=TRUE)
}
```

## By fusion, by trial 

Tabulate age group, sex, protocol, and outcome by fusion group and trial. 

```{r}
if (requireNamespace("tableone")) {
  CreateTableOne(data=subset(covs), strata=c("FusionGroup","Protocol"), 
                 includeNA=TRUE)
}
```
